\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-31 trunk/arch.tex
- Updated europa solve algorithm so I can refer to it - updates in the t-rex section

{\em\color{red} Following is an attempt to present TREX as a big
  deliberation machine}
The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:
\begin{itemize}
\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.
\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.
\end{itemize}
Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple
ticks for a reactor to produce a plan that is both complete and
valid. when such plan is identified its {\em external} part is then
dispatched to the corresponding reactors which in turn can start their
deliberation \texttt{step}s in order to satisfy these new
objectives. as a consequence physical execution will occur as these
goals and planning propagate down in the reactor graph until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal in a physical command
execution. 
{\em\color{red} end of the attempt; need to rework the paragraphs
  below accordingly}

The overal execution cycle of the agent is emerging through these
interaction and the processing that occurs in each reactors in order
to process them. New observation occurring at every tick inform each
reactor of the current state of the world and impose them to identify
how it impact their state (and potentially what they planned for the
future). Goals they receive will trigger in the reactor the need to
identify a strategy (or plan) that will allow them to reach these new
states knowing their current state deduced from synchronization.When
such a plan has been identified, the {\em external}  part is then
dispatched to other reactors that can do the same. This flow and focus
around state and goals is perfectly fitting with automated planning
and even more so to the Europa framework. As an illustration, in 
 Fig. \ref{fig:agent}, only \textsf{Shore comm.} and \textsf{Vehicle}
 reactors are specialized C++ code (acting as interface to external
 programs). All the other reactors are based on the europa framework
 differentiated only by the domain model they manipulate and som
 heuristic configuration parameters. In the following sections we are
 going to focus specifically ion this kind of reactor to illustrate
 how  a formal planning framework can be used in our architecture to
 both track state evolutions and plan for future objectives while
 remaining efficient enough to do these tasks in a real-time control 
loop.

@@

In T-REX architecture reactors have two core functions;
synchronization where the main focus is to track and refine the
evolution of the state of the world at any single tick, and
deliberation which focus on providing a plan to the agent in order to
complete the reactor {\em Internal} goals. These two processes are
intertwined in the sense that one influences the outcome of the
other. They are represented as two europa solvers modifying the same
plan database and for which the execution is managed by the T-REX 
framework as time advance during the execution of the system:
\begin{enumerate}
\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.
\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.
\end{enumerate}

While synchronization is assumed to be fast enough that it can be
resolved in less than a tick. Deliberation on the other hand can take
several ticks reflecting that automated planning can take time
depending on how far in the future the planner is looking ahead and
the complexity of the model. T-REX architecture, and more specifically
the europa based reactor in our case, takes this into account by
considering synchronization and deliberation as two concurrent tasks
with their own timing requirements and priority. 
While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}. Fig. \ref{fig:tick-exec}
shows a conceptual view of these two tasks.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

In  this section we present how both processes are implemented using
the europa framework. We first present each of them independently by
looking to their specific focus. We then develop how both processes
are not only interleaved (as deliberation can take several tick while
synchronization eneds to happen at every single tick) but also discuss
on how each of these is impacting the other and how this intertwining 
of the two result on an emergent behavior of the reactor by allowing
the system to efficiently adapts its plan that can then be dispatched
as objectives for the reactors it depends on in a seamless flow.

{\em fpy: I still wonder whether I should start with planning assuming
  first that the initial state is known and then develop
  synchronization and how it ties state information in to the plan or
  the other way around as it is now}

>>>>

The choice of representation of state variables, observations and
goals is directly derieved to the timleines and tokens from europa. 
such representation is ich enough to allow the excahnge of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time 
(such as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:
\begin{itemize}
\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.
\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.
\end{itemize}
Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple
ticks for a reactor to produce a plan that is both complete and
valid. when such plan is identified its {\em external} part is then
dispatched to the corresponding reactors which in turn can start their
deliberation \texttt{step}s in order to satisfy these new
objectives. As a consequence physical execution will occur as these
goals and planning propagate down in the reactor graph until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal in a physical command
execution.

The way the agent interleave planning \texttt{step}s and
synchronization is illustrated in Figure \ref{fig:tick-exec}.
Synchronization happens at every single tick and will interrupt
planning allowing this reactor to identify its current state that can
propagate through the reactor graph in order to ensure t aconsitent
view of the state of the world between all the reactors.  
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}
This view of the reactor has two parallel tasks aloow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The europa reactor is our implementation inside T-REX of a reactor
that usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea we of T-REX we
described above to implement a reactor for which the sole purpse is to
deliberate as time advance. The choice of europa as the planning
framework was directed by the fact thta this framework provide  a rich
and highly configurable framework to be able to infer around a rich
temporal domain.

This reactor is implemented following the idea that in T-REX, the
whole execution cycle is two deliberation processes. Both of these
processes are implemented as two europa solvers modifying the same
plan database and for which the execution is managed by the T-REX 
framework as time advance during the execution of the system:
\begin{enumerate}
\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.
\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.
\end{enumerate}
While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}.

An important challenge for embedding automayted planning i na situated
agent as always been to go around the asumptions made by most classical
planners. Specifically one assumption restrict the problem to ``offline
planning''  defined in \cite{ghallab04} as follow:
\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{In this book $\Sigma$ identifies the world modeled
    by the plan domain} while it is planning; it plans for the given
  initial and goal states regardless of the current dynamics, if any.
\end{quotation}
This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is righe before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In 
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by aloowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}

Looking back at Fig. \ref{fig:tick-exec}, one can already see that
this asumtpion is problematic in our architecture. Indeed
synchronization that tracks state evolution can occur several times 
during the reactor deliberation process. By making synchronization and
deliberation share the same plan we provide a solution that relax this
assumption. In the following sections we present how synchronization
and deliberation are implemented independently first to finally show
how their interaction through the plan helps relax this assumption by
making the planning fully integrated to the reactor execution cycle
while potentially allowing for more informed planning.
